---
title: "Basic inferential data analysis"
author: "Palani"
date: "Sunday, June 05, 2016"
output: word_document
---

data(ToothGrowth)

The first 5 rows are
```{r eval=TRUE,echo=FALSE}
head(ToothGrowth)
```

The dimension of the data is
```{r echo=FALSE,eval=TRUE}
dim(ToothGrowth)
```

#### Exploratory Analysis

*Conditioning Plot*  
This plot compares two factor variables with the one numeric variable .And the plot clearly explains that there is a difference in dosage levels and delivery method  with respect to length of teeth in each of 10 guinea pigs.

```{r echo=TRUE,eval=TRUE}
coplot(len ~ dose | supp, data = ToothGrowth, panel = panel.smooth,
       xlab = "ToothGrowth data: length vs dose, given type of supplement")
```

Similar analysis using `Box` plot.

```{r,echo=FALSE,eval=TRUE}
library(ggplot2)
ToothGrowth$Combination <- interaction(ToothGrowth$supp, ToothGrowth$dose)
g<-ggplot(aes(y =len, x = Combination), data = ToothGrowth) + geom_boxplot()
g+ggtitle("Interaction Box Plot")+theme(axis.text = element_text(colour = "blue"))+
  theme(axis.text=element_text(size=12),axis.title=element_text(size=20,face="bold"))+
  theme(plot.title = element_text(lineheight=.8, face="bold"))
```



**3-Way Frequency Table Analysis**

This numerical analysis clearly shows there is a clear difference between two factors and numeric variable length of teeth.

```{r echo=TRUE,eval=TRUE}
mytable <- xtabs(ToothGrowth$len~ToothGrowth$supp+ToothGrowth$dose, data=ToothGrowth)
ftable(mytable) # print table 
```

*Mean anlsysis by group*
There is a clear difference in mean between delivery method and length of teeth as well as dosage levels and length of teeth.

```{r echo=TRUE,eval=TRUE}
tapply(ToothGrowth$len,list(ToothGrowth$supp,ToothGrowth$dose),mean)
```

#### Assumptions 
**Considering the following assuptions are valid before conducting t.test**

 + The t interval technically assumes that the data are iid normal  
 + The second assumption made is that of a simple random sample, that the data is collected from a representative, randomly selected portion of the total population.
 + The third assumption is that the data, when plotted, results in a normal distribution, bell-shaped distribution curve.
 + The fourth assumption is that a reasonably large sample size is used. A larger sample size means that the distribution of results should approach a normal bell-shaped curve.
 + The final assumption is homogeneity of variance. Homogeneous, or equal, variance exists when the standard deviations of samples are approximately equal.

### Inferential Analysis  
*Setting Hypothesis*  

**H0=mu0=No Difference between `Delivery` methods.**  
**Ha=mua=There is a difference betwwen `Delivery` methods.**  

*Evaluating hypothesis through estimating `Confidence Intervals`*  
*t,tests are used for constructing confidence interval*  

Analysing variance between groups for setting variance parameter in t.test procedure.


```{r echo=TRUE,eval=TRUE}
tapply(ToothGrowth$len,list(ToothGrowth$supp,ToothGrowth$dose),var)
```

The table shows there is non constant variance across delivery methods and length of teeth.

Evaluating hypothsis test through t.test

```{r,echo=TRUE,eval=TRUE}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth)$conf
```

The confidence interval contains 0 hence there is no signifigant difference in delivery methods.

Estimating `p` value

```{r echo=TRUE,eval=TRUE}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth)$p.val
```

The p value is greater than alpha level 0.05 hence we fail to reject null hypothesis.

### Inference Analysis 2

**H0=mu0=No Difference between `Dosage` levels.**  
**Ha=mua=There is a difference betwwen `Dosage` levels.**

This time comparing dosage levels and length of teeth growth.

Unequal variance T test comparing 3 dosage levels with length of teeh.  

```{r echo=TRUE,eval=TRUE}
wideCW14 <- subset(ToothGrowth, dose %in% c(0.5, 1))
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = wideCW14)$conf
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = wideCW14)$p.val
```


```{r echo=TRUE,eval=TRUE}
wideCW14 <- subset(ToothGrowth, dose %in% c(0.5, 2))
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = wideCW14)$conf
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = wideCW14)$p.val
```

```{r echo=TRUE,eval=TRUE}
wideCW14 <- subset(ToothGrowth, dose %in% c(1, 2))
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = wideCW14)$conf
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = wideCW14)$p.val
```


**Interpretation**  
The above all the three analysis shows there is a difference between all the dosage levels and length of teeth in pigs.
The reason for confirming for the significant difference is the confidence interval does not contain zero and the p value is less than alpha.Hence we reject `null` hypothesis.  
And the confidence interval describes the percentage of intervals that would cover the parameter being estimated if we were to repeat the experiment over and over.

#### Conclusion  

**In this inference analysis , followings inferences are  concluded ,**  

  + There is no difference between delivery methods and length of teeth pig.  
  + There is a significant difference between doasge levels and length of teeth pig.  
 
 
 










